### 2017-04-04 Doc/classicthesis/Chapters/Chapter02.tex

%*****************************************
\chapter{State of the art}\label{ch:stateoftheart}
%*****************************************

In this chapter we present the two dimensions of this project: the \ac{IoT} development state and an introduction to IBM's privacy-preserving solution, Idemix.

\section{Internet of Things}

\section{Idemix}