### 2017-04-06 Doc/classicthesis/Chapters/Chapter03.tex
vendehumos

%************************************************
\chapter{Objectives and Methodology}\label{ch:objandmeth}
%************************************************

%TODO resumir organización del capítulos

\section{Project description}

\section{Working methodology}

\section{Development environment}

% C as pure as possible aiming to lower powerful devices without C++ support or std libraries
% CMake for portability and easy to config cross-compilation
% Java for P2ABCE
% Docker for building environments, both maven and LEDE

% HW: Omega2 with LEDE
%	  Raspberry Pi 3 with Raspbian