### 2017-05-19 Doc/classicthesis/Chapters/Chapter0A.tex
fixes + organización

%********************************************************************
% Appendix
%*******************************************************
% If problems with the headers: get headings in appendix etc. right
%\markboth{\spacedlowsmallcaps{Appendix}}{\spacedlowsmallcaps{Appendix}}
\chapter{Test: APDU Commands exchanged}\label{ch:resultsdiagrams}

% TODO
In this appendix we provide 3 sequence diagrams highlighting the APDU Commands exchanged during our testbed. The first one is the setup of the IoT smart card, storing the system parameters needed to work in the deployed P2ABCE system. The second one is the issuance of the credential, divided in the three REST calls needed in the delegation. The last one is the proving for the Presentation Policy from the Verifier, using two REST calls.

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=0.9\linewidth]{gfx/APDUsInitIoTSC}
	\end{center}
	\caption{Init IoT Smart Card APDU Commands exchanged.}
	\label{fig:APDUsInitIoTSC}
\end{figure}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=0.78\linewidth]{gfx/IssuanceAPDUs}
	\end{center}
	\caption{Issuance APDU Commands.}
	\label{fig:IssuanceAPDUs}
\end{figure}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/APDUsProving}
	\end{center}
	\caption{Proving APDU Commands.}
	\label{fig:APDUsProving}
\end{figure}

\chapter{Guide on how to continue the project development}