### 2017-05-19 Doc/classicthesis/Chapters/Design.tex

%************************************************
\chapter{Design and implementation}\label{ch:design}
%************************************************

%TODO resumir organización del capítulo