### 2017-06-05 Doc/classicthesis/Chapters/Design.tex
impl avance

%************************************************
\chapter{Design and implementation}\label{ch:design}
%************************************************

%TODO resumir organización del capítulo

% Estructura:
% +Design
%  -Architecture design: delegation, computation offloading, duality User/SC + Engine
%

In this chapter we will describe the process for defining how a constrained IoT device may be integrated in a system like P2ABCE. We will also describe the PoC implementation carried out to test a realistic deployment.

\section{Design}

In this section we will define how an IoT device may be integrated in the P2ABCE architecture, being totally compatible with any other system using P2ABCE, addressing the power and memory constrains many IoT devices face.

We decided to use P2ABCE with the Idemix as its Engine, because it is officially supported by the Idemix Library, it has the most up-to-date implementation, as we saw in the state of the art, and adds capabilities to Idemix like the Presentation Policies, or interoperability with U-Prove, not available without the P2ABCE project.

Our main goal is to make an IoT device capable to act as a User or Verifier in the P2ABCE architecture. For this, the device should be able to \textbf{communicate} with the Verifier or Prover with which it is interacting, manage the P2ABCE complex \textbf{XML schemas} transmitted, and perform the \textbf{cryptographic operations} required.

The communication between actors depends on each IoT scenario, it can be achieved with many existing standard solutions, e.g. an IP network, a Bluetooth M2M connection, RF communication, etc.

Our real concerns are, on one side, parsing the XML data, based on the P2ABCE's XML schema, that specifies the data artifacts created and exchanged during the issuance, presentation, revocation and inspection of pABCs; and on the other side, the cryptographic operations, that involve the use of secret keys, stored privately in the IoT device.

After the analysis done in the previous sections to the P2ABCE architecture, emphasizing that the logic of smart cards gathers the cryptographic operations independently from how the data is exchanged between P2ABCE actors.

Using the \textit{computation offloading} technique to our scenario, our design consists on implementing the smart card logic inside the IoT device, keeping secure our master key and credentials, and for the rest of the P2ABCE system, if the device can not run the complete Engine, it may delegate to a server running it, indicating how to send APDU Commands to the \textit{IoT smart card}.

Even in the case we were to implement all P2ABCE inside an IoT device, we would have to implement the support for software smart cards, to keep the secret inside the IoT device. Therefore, we can begin implementing the smart card logic inside the IoT device, and later, if the device resources admit it, other components of the P2ABCE project.

\hfil

Computation offloading is not new to IoT deployments. For example, IPv6 involves managing 128 bits per address and other headers, and many IoT scenes only need to communicate inside a private network, making only the last 64 bits in an address relevant. To reduce that overhead, instead of IPv6 they use 6LoWPAN to compress packets and use smaller address sizes. To communicate a 6LoWPAN with the Internet or other networks, the IoT devices delegate the networking workload on a proxy that can manage the 6LoWPAN and IPv6 stacks. In the scope of consumer devices, smart bands or watches can install applications, bit many of them delegate on the user's phone to accomplish their task.

Therefore, the IoT device now has a \textbf{duality} in its functions, because it is the User that starts any interaction with other actors, and it's also the smart card that a P2ABCE server must ask for cryptographic operations. It can also be seen as a \textbf{double delegation}. The IoT device delegates on the external P2ABCE server to manage the protocol, and the P2ABCE server delegates on the IoT, acting now as a smart card, for the cryptography.

\subsection{System architecture}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/P2ABCE-IoT-bw}
	\end{center}
	\caption{IoT in P2ABCE Architecture.}
	\label{fig:P2ABCE-IoT}
\end{figure}

%\begin{figure}[bth]
%	\begin{center}
%		\includegraphics[width=\linewidth]{gfx/P2ABCE-IoT-color}
%	\end{center}
%	\caption{IoT in P2ABCE deployment diagram.}
%	\label{fig:P2ABCE-IoT-color}
%\end{figure}

\hfil

The system will be compounded by the IoT device, the P2ABCE delegation server and the third party P2ABCE actors.

\begin{itemize}

	\item \textbf{IoT device}

	In \autoref{fig:P2ABCE-IoT} the IoT device is represented with two interfaces, physical or virtual. One allows external communications to other machines, including other P2ABCE actors, that could be on the Internet, a corporate LAN, a M2M overlay network, etc. Through this interface, the P2ABCE XML messages are exchanged as in any other deployment. This allows an IoT device to interact with other actors without special adaptations to the protocol. The other interface allows a secure communication with the delegation server. Both the delegation messages and the APDU Dialogue are transmitted over this interface, making it a point of attack to the system, and we will talk about its security in the delegation process.

	The scheme also shows the \textit{P2ABCE IoT Toolkit}. This piece of software includes the IoT Smart Card, and the API for other processes that want to use the P2ABCE system.

	The IoT Smart Card is the implementation of a software smart card, listens for APDU Commands from the secure interface and stores securely the credentials and private keys within the device's memory.

	The P2ABCE API is an interface for other processes that wish to use the private-preserving environment of P2ABCE. It provides access to every operation available, hiding the delegation process. In the future, if for example the Verification Service is implemented for the IoT device, i.e., there's no need to delegate to other machine to act as a Verifier, then any program using the API won't need to change anything, the toolkit conceals the transition from delegating to native execution.

	\item \textbf{P2ABCE actors}

	If we recall from \autoref{analysisP2ABCE}, the possible roles in the system were the Issuer, the User, the Verifier, the Revocation Authority and the Inspector. All of them use the P2ABCE XML schema in the specification to communicate to each other. Any third party actor will be unaware of the fact that the device is a constrained IoT device that delegates on the P2ABCE server.

	\item \textbf{P2ABCE Delegation Server}

	The machine in charge of receiving authorized IoT devices' commands to parse the XML files exchanged and orchestrate the cryptographic operations the IoT smart card must perform.

\end{itemize}

\hfil

\begin{flushleft}
	\textbf{Delegation process}
\end{flushleft}

Here we describe the computation offloading carried out by the IoT device.  In \autoref{fig:DelegationProving} we show an example of the IoT acting as a User, Proving a Presentation Policy to a third party Verifier.

\begin{enumerate}
	\item Communication with P2ABCE actor.

	The IoT device starts an interaction with another actor, e.g. an Issuer or Verifier, receiving a P2ABCE XML file.

	\item Delegation to the P2ABCE Server.

	Depending on what role the IoT device is acting as, it will delegate in the corresponding service, e.g. User Service. The delegation message must include the XML file, and any parameter required to accomplish the task, like the information on how to communicate with the IoT smart card (listening port, security challenge, etc.).

	\item APDU Dialogue (if necessary).

	The server may need to send APDU Commands to the IoT smart card to read the credential information or perform cryptographic operations involving private keys.

	\item Server response.

	The server may return a status code or a XML file if the first one required an answer from the IoT device, in which case, it will send as response to the third party actor, resuming the communication.

\end{enumerate}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/UML/DelegationProving}
	\end{center}
	\caption{IoT Delegation in P2ABCE for Proving.}
	\label{fig:DelegationProving}
\end{figure}

Transmissions over the \textit{Server-IoT} channel must be secured in order to avoid attacks like: impersonate the P2ABCE delegation server, having access to the IoT smart card sending the APDU Commands the attacker wishes; delegate as a device on the server but giving the parameters of another device, making the delegation server send the APDU Commands to a victim IoT smart card.

We could use a corporative PKI to issue certificates to the server and devices and configure policies for access control; design a challenge-response system combined with the smart card PIN, like a password and TOTP\footnote{Time-based One-time Password} in a 2FA\footnote{Two-factor authentication} login. We also could connect physically the delegation service through RS-232 serial to the IoT device, securing both physically as we would do with the IoT device on its own, isolating the delegation system from any network attack. This last idea is an approximation to the Arduino Yún\footnote{\url{https://www.arduino.cc/en/Main/ArduinoBoardYun}}, a development board that integrates two microcontrollers, one a typical Arduino with very low resources, and another one running a fork of OpenWrt. The Arduino microcontroller can control the terminal of the more powerful one, using the serial pins as commented before.

As we can see, there are many state of the art solutions for all this threads, therefore, we can assume a secure channel without mentioning a specific solution, providing freedom to choose the most fitting one in a real deployment.

\paragraph{Notes for more constrained devices}

Our architecture is designed for devices that could in a future run a reimplemented version of P2ABCE, that means, the devices could perform more tasks than only running the smart card software and their main purpose process, e.g. recollecting sensor data. But if our target devices are so constrained that can barely run the smart card, they may not be able to handle the XML files because of memory restrictions, like a MSP430\footnote{\url{http://www.ti.com/lsds/ti/microcontrollers-16-bit-32-bit/msp/overview.page}} running Contiki-OS, the microcontroller has between hundred of bytes to tens of kilobytes of memory, making impossible to store multiple XML files in the size range of tens of kilobytes.

In these cases, the delegation in the server goes a step forward, making the server a proxy to communicate with other P2ABCE actors, and the IoT device only acts as a smart card. The IoT device would still act as the User, or any other P2ABCE, role because it orchestrates when and how an interaction with other actor is executed, but the communications would be between the proxy and the third party actor.

\section{Proof of Concept Implementation}
% +Implementation:
% *System:
% 	-Raspbian OS: distro de debian para Raspberry Pi, que se describirá en los tests
% 	-LEDE: ash terminal (shell command interpreter), POSIX, typical hardware, libraries with no hardware aceleration vs smart card MULTOS,
%
% *Delegation
%  -REST as current delegation protocol: curl script bash
%  -BIOSC as current APDU transmission protocol: no security -> no overload.
% *Smart Card
%  
%  -IoT Smart Card software design and implementation notes.
%  -Sequence diagrams. TODO

In this section we present the first PoC implementation, introducing the IoT system where we are going to work, then we will describe the delegation protocols, one for the computation offloading of the IoT device on the P2ABCE server, and another for the transmission of APDU Commands, and finally, we will describe the IoT smart card implementation.

\subsection{IoT system}

We will develop our PoC in Linux based systems in order to avoid complications with firmware specific issues that we are not familiar with. Even so, the linux systems used are aimed for the IoT environment and serve as a starting point for future implementations in more constrained devices or different systems.

In our delegation server we will run Raspbian OS, a distribution based on Debian for the Raspberry Pi. We will talk more about the hardware specifications in the benchmark chapter. We chose this system because it has a great package repository support, including the Java Runtime Environment, needed to run the P2ABCE Services.

For our IoT device, we will use LEDE, Linux Embedded Development Environment, a distribution born from OpenWrt, aimed for routers and embedded chips with low resources requirements. It offers the \texttt{ash} command interpreter, or \texttt{shell}, and the \texttt{opkg} package manager, that allows an easy installation of some libraries and tools needed during the early development.

\subsection{PoC Delegation}

As we explained in the design section, the delegation has two steps, the IoT device calling the P2ABCE server to offload the parsing of the XML data, and the P2ABCE server sending APDU Commands to the IoT smart card in the device.

\subsubsection{PoC Delegation to the P2ABCE Server}

Currently P2ABCE offers multiple REST web services to run different roles in P2ABCE system: User Service, Issuer Service, Verification Service, etc. Any third party application that integrates P2ABCE with their system can make use of these services or implement the functionality using the library written in Java, the tool that the REST services actually use.

Our PoC machine, the Omega2, can make REST calls easily with the \texttt{curl} command, but other devices may use \ac{CoAP}, but in that case, the P2ABCE REST services should be adapted to offer CoAP support. The commands needed to delegate to the P2ABCE server would be the same that those defined now to operate with the REST services, but with the modifications needed to pass the parameters on how to communicate with the IoT smart card.

In this PoC the only P2ABCE Service that needed to be modified was the User Service. We added the following REST call

\begin{center}
	\textit{/initIoTsmartcard/{issuerParametersUid}?host=\&port=}
\end{center}

where we communicate the P2ABCE server that an IoT Smart Card is accessible via \textit{host} and \textit{port}. Then a new \textit{HardwareSmartcard} object is stored in the P2ABC Engine, but instead of the \textit{javax.smartcardio} Oracle's \textit{CardTerminal} implementation, we use our own \textit{IoTsmartcardio} implementation for the \textit{HardwareSmartcard} constructor, that we will discuss in the following section.

\hfil

In a real deployment, we would offer a full library with an API for other processes that want to use the P2ABC Engine, which automatizes the boot of the IoT smart card and handles the security policies, like the one we presented in the deployment diagram \ref{fig:P2ABCE-IoT} of the \textit{P2ABCE IoT Toolkit}.

Now we show some examples of the PoC \texttt{curl} commands, run in the device's terminal, or from a shell script for automation:

\begin{lstlisting}[language=bash]
$ curl -X POST --header 'Content-Type: text/xml' "http://$DelegationServerIP:9200/user/initIoTsmartcard/http%3A%2F%2Fticketcompany%2FMyFavoriteSoccerTeam%2Fissuance%3Aidemix?host=192.168.3.1&port=8888"

$ curl -X POST --header 'Content-Type: text/xml' -d @firstIssuanceMessage.xml "http://$DelegationServerIP:9200/user/issuanceProtocolStep/" > issuanceReturn.xml
\end{lstlisting}

The IoT device will only manage XML as data files to exchange between the third party actor and the P2ABCE delegation service, attaching them in the body of the REST call. In the parameters of the first call we can see the IP and port where the IoT smart card will be listening.

\subsubsection{APDU Dialogue Transmission}

To transmit the APDU messages in our PoC we use a simple protocol, that we will refer as BIOSC (Basic Input Output Smart Card), consisting in one first byte for the instruction:

\hfil

\begin{tabular}{|c|l|}
	\hline
	\texttt{0x01} & APDU Command or Response: Read a 2 bytes for the\\& length and then as many bytes as said length. \\
	\hline
	\texttt{0xff} & Finish connection: close the current TCP socket and\\& end the IoT smart card execution. \\
	\hline
\end{tabular}

\hfil

In the first instruction, we read two header bytes with the length of the APDU Command or Response to receive, followed by said APDU bytes. The message is sent over TCP for a reliable transmission (concept of session, packet retransmission, reordering, etc.). The second instruction serves for closing the TCP socket in the IoT device when the smart card is no longer needed.

We lack any security (authentication or authorization) that a real system should implement. It is vital to authenticate the delegation service, to authorize it to make APDU Commands, and the same with the IoT device, to prevent attacks. But using this method for the transmission of the APDU messages, with only 3 bytes of overhead, helps us in the  benchmarks to measure the real performance of the system.

\hfil

The implementation of this simple protocol is done in Java and C, one for the P2ABCE server and the other one for the IoT smart card.

% TODO : IoTsmartcardio
% TODO : BIOSC.c
%TODO : optional: diagrama secuencia de P2ABCE, pero si eso solo de la reimplementación de javax.smartcardio y su uso por HardwareSmartcard

\subsection{IoT Smart Card Implementation}

After many design decisions in the process to adapt the original ABC4Trust Card Lite code to pure C, working on a MIPS architecture, in this section, we present the current \ac{PoC} code, most important decisions and execution workflow in the form of sequence diagrams.

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/UML/sequenceBIOSC}
	\end{center}
	\caption{IoT Smart Card Sequence Diagram.}
	\label{fig:sequenceBIOSC}
\end{figure}

\hfil

%%%%%%%%%%%%%%%%%%%%%%%

%% DRAFT %%

\subsection{Code structure}

We divide the project in three different sections with the objective of enhancing maintainability, improving future changes, ports, fixes, etc.

The first section is what could be called as the core of the smart card, the second one the interface for those tools the core needs and may depend on the target's platform, and finally, the third party libraries, that may be empty if the interfaces implementation doesn't need any.

%\begin{figure}[bth]
%	\begin{center}
%		\includegraphics[width=\linewidth]{gfx/IoTCScomponents-color}
%	\end{center}
%	\caption{IoT Smart Card Code Structure.}
%	\label{fig:IoTCScomponents-color}
%\end{figure}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/IoTCScomponents-bw}
	\end{center}
	\caption{IoT Smart Card Code Structure.}
	\label{fig:IoTCScomponents-bw}
\end{figure}

\hfil

\paragraph{Core smart card}

The smart card logic lies in this section, the concepts of APDU Commands, what instructions are defined in P2ABCE smart cards, and how to process them and generate proper APDU Responses.

If the APDU protocol defined for P2ABCE smart cards changed, with the consequence of having to update all smart cards in use, the changes to adapt the IoT smart card should be applied here. This also implies that if the P2ABCE Java code changes their \texttt{Smartcard} interface, as long as it is backwards compatible with existing smart cards, the IoT smart card will still work as intended.

Most of the original ABC4Trust Card's code has been reused in this section, like all the variables and custom data type definitions, or the APDU Commands handling.

However, the ABC4Trust's code heavily depended on the MULTOS C-API for the input/output of data, modular arithmetic, and AES128 and SHA256 cryptographic operations.

A characteristic of MULTOS C-API is that every function name starts with ``\texttt{multos}'', so we implemented, following the C-API specification, the functions used, but changing their names from \texttt{multosFoo()} to \texttt{mFoo()} for readability and to emphasize that they were no longer MULTOS functions.

The \texttt{main.h} file, from the original project, also implemented equivalent functions to some available in \texttt{multos.h}. We replaced them for the standard functions in the C-API, checking that the MEL code did the same as the documentation specified. This last detail is very important, because in \citep{vullers2013efficient} they noted that \texttt{ModularExponentiation} function dos not accept exponents larger than the modulus size, and they added \texttt{SpecialModularExponentiation}, dividing the exponentiation in two that MULTOS could perform. After this analysis, when those particularities appeared, we decided to adapt the reimplementation of the C-API.

Future changes in the code may refactor it so there's no longer need for the MULTOS framework functions.

\paragraph{Interfaces}

To implement MULTOS functions, we needed to use some libraries, so we defined a facade to isolate the implementation of the core smart card from our different options, that could vary depending on the hardware or the system used by the IoT device.

The use of a facade lets us, for example, change the implementation of modular arithmetic with a hardware optimized version, or a future more lightweight library, or our very own software implementation using the same data types that the core uses, minimizing the data transformations needed.

Taking a step forward, we make the core smart card totally independent of any library, only on our interfaces. This means that typical C libraries, like the standard \textit{stdlib.h}, or  \textit{string.h} are also behind the facade, in case some IoT system doesn't support them. The main goal we go after with this decision is that future developers adapting the code to a specific platform need to make no change to the \textit{core smart card}'s code, only to the interfaces implementation.

\paragraph{External utilities}
% Podría haber implementaciones de librerías criptográficas en ensamblador de la plataforma, se incluyen aquí y la interfaz lo usa al implementar

If the IoT system offers well tested libraries that could aid in the interfaces implementation, or we simply found a pure C implementation for the task, these third party libraries belong to this section.

In our PoC, we use two ANSI C libraries, for base64 and JSON, and two shared libraries available in as packages in LEDE, GMPLib and OpenSSL. The last two libraries offer more functionality than we need, hence, it's desired in a production code to implement \textit{Modular Arithmetic} and \textit{Cryptography} interfaces with more lightweight alternatives.

For example, Atmel's ATAES132A \citep{ATAES132A}
\marginpar{\includegraphics[width=0.9\linewidth]{gfx/atmel} \\ Atmel's cryptography chips.}
offers a serial chip for secure key storage, AES128 execution and random number generation. Another serial chip like ESP8266 offers WiFi connectivity, typically used with Arduino, and can also perform AES encryption. For random number generation, a technique used with Contiki devices is to read from sensors aleatory data and use it as seed. All these alternatives depend on the target device, but are all valid. The \textit{interfaces} and \textit{external utilities} sections allow for a clean and fast port of the code.

% PoC BIOSC : TCP sockets, JSON serialization

% PoC IoT javax.smartcardio : protocolo de 2 bytes + APDU

\subsection{PoC Workflow}

% 1. Boot: deserialize sc status
% 2. Listening: open tcp socket
% 3. The IoT device delegates on the P2ABCE for some task and sends the IP and Port of the IoT Smart Card.
% 4. An APDU Command arrives : read in 2 bytes the length of the apdu and

In a real deployment, we would offer a full library with an API to other processes to delegate on the P2ABC Engine, that automatizes the listening and security, that we presented in the deployment diagram \textbf{TODO:ref-img} as the \textit{P2ABCE IoT Toolkit}.

But to test the IoT smart card, we use the \textit{curl} command for REST delegation calls and the BIOSC (Basic Input Output Smart Card) for APDU communication.

To transmit the APDU messages in our PoC we use a simple protocol, consisting in one first byte for the instruction: receive an APDU Command or close the connection. In the first case, then we read two header bytes with the length of the APDU Command or Response to receive, followed by those APDU bytes. The message is sent over TCP for a reliable transmission (concept of session, packet retransmission, reordering, etc.).

We lack of any security (authentication and authorization) that a real system should implement. It is vital to authenticate the delegation service, to authorize it to make APDU Commands, and the same with the IoT device, to prevent attacks. This belongs to \textit{usual} security, as we already said, and for that reason it's not in the PoC.

\paragraph{APDU transmission with BIOSC}

Before sending the REST message, the Omega2 must be listening in the specified port. At start up, BIOSC reads from a JSON file the status of the smart card (credentials in the BLOB, private keys, PIN and PUK codes, etc.), then opens the TCP socket and listens in a loop.

When the delegation server sends an APDU Command following the simple protocol, BIOSC stores in a byte array the APDU Command bytes and calls the \textit{handle\_APDU} function in \textit{Core Smart Card}.

The \textit{APDU\_handler} will parse the APDU bytes and check both CLA and INS bytes. For each possible INS the is a function that must always finish calling either \textit{mExit()} or \textit{output\_large\_data()}.

The \textit{mExit} function is the reimplementation of MULTOS C-API exit functionality, that finishes the application execution returning to the MULTOS OS, that will send the APDU Response bytes to the terminal. This led in the original ABC4Trust Card's code to some tricky situations. Imagine a function that ends with an if-else expression. Many times we save a line writing something like

\begin{verbatim}
if(condition)
return a;
return b;
\end{verbatim}

instead of the complete

\begin{verbatim}
if(condition)
return a;
else
return b;
\end{verbatim}

This is a swift example of what the \textit{multosExit} function led to, because it is called in both the \textit{switch-case} processing the APDU instruction, and the \textit{Subroutines}. For example, the PIN code check was programmed in a way every reason to fail finished the application execution, with different error codes. Another example, checking if a credential is stored, if not, fail and exit inside the \textit{readBlob} subroutine.

In our standard architecture machine, if we call a subroutine, it must end, and return the control to the handling function, and so the \textit{APDU\_handler} must return to the listening loop of BIOSC.

Having all that in count, the \textit{mExit} function is not implemented as the documentation specifies, but only saves the smart card status serializing it in the JSON file again, and once it's saved, parses the APDU Response and sends it to the socket.

The \textit{output\_large\_data} function is a tool used in MULTOS smart cards that don't support Extended APDUs. The APDU Command handler saves in a buffer the data to send, and its size. Then, \textit{output\_large\_data} will manage the buffer expecting future GET RESPONSE commands from the P2ABCE server.

A future change to the project may include support for Extended APDUs, avoiding the use of multiple GET RESPONSE. To do this, the \textit{output\_large\_data} function and the \textit{IoTsmartcardio} implementation of \textit{javax.smartcardio} must support the feature.