### 2017-06-19 Doc/classicthesis/FrontBackmatter/Abstract.tex
ieee

%*******************************************************
% Abstract
%*******************************************************
%\renewcommand{\abstractname}{Abstract}
\pdfbookmark[1]{Abstract}{Abstract}
\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Abstract}

In this paper we design a solution to integrate IBM's Identity Mixer into the IoT ecosystem, integrating its privacy-preserving capabilities thanks to the use of Zero-Knowledge Proof cryptography. To validate our design, we implemented a functional Proof of Concept written in C for IoT devices. The applications of such solution promise to improve both the security and privacy of IoT, which in recent years have proved to be compromised.

\vfill

%\begin{otherlanguage}{ngerman}
%\pdfbookmark[1]{Zusammenfassung}{Zusammenfassung}
%\chapter*{Zusammenfassung}
%Kurze Zusammenfassung des Inhaltes in deutscher Sprache\dots 
%\end{otherlanguage}

\endgroup

\vfill